A Fredholm operator approach to eigenvalue problems in elasticity and planar Stokes flow

### 2019-01-16 doc/draft-01/02prelim.tex
add matlab files (things are still a bit rough) also added mathematica notebooks that are less crucial

Let $L_x$ denote a linear differential operator. A fundamental
solution $G(\xx,\yy)$ of $L_x$ satisfies the equation
$L_x G(\xx,\yy) = \delta_y(\xx)$ in the distributional sense, i.e.
for sufficiently smooth $f$
\begin{equation}
  L_x \int_{\R^2} G(\xx,\yy) f(\yy) \, dy = f(\xx) \; .
\end{equation}
We consider here
``free-space'' Green's functions, i.e. fundamental solutions which satisfy
some natural growth or decay conditions as $|\xx-\yy| \to \infty$.
The Green's function of the oscillatory biharmonic equation
\cref{eq:buck1} is given by 
\begin{equation}
  \Gbh(\xx,\yy) = \frac{1}{k^2}
  \left (\frac{1}{2\pi} \log |\xx-\yy| +
  \frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \right ) \, , 
\end{equation}
where $k$ is the Helmholtz paramater in the oscillatory biharmonic equation,
and $H_{0}^{1}(r)$ is the Hankel function of the first kind of order zero.
Note that this is a scaled difference of the Green's function for
Laplace, i.e.

@@

\subsection{A Farkas-style integral representation}

In \cite{Farkas89}, Farkas derived integral equation formulations
for the solution of the Dirichlet problem for the biharmonic equation ($k=0$). 
In this formulation, the solution 
$w$ is represented as a sum of two layer potentials
\begin{equation}
  w(\xx) = \int_\Gamma \Kfark_1 (\xx,\yy) \sigma_1(\yy) \, dS(\yy)
  + \int_\Gamma \Kfark_2(\xx,\yy) \sigma_2(\yy) \, dS(\yy) \; ,
  \label{eq:farkRep}
\end{equation}
where $\sigma_{1}$, and $\sigma_{2}$ are unknown densities, and the kernels
$K_{1}^{F}$, and $K_{2}^{F}$ are linear combinations of derivatives of
the biharmonic Green's function and appropriately chosen to get second kind integral equations
for the unknown densities $\sigma_{1}$ and $\sigma_{2}$ on imposing the boundary conditions.

In \cite{jiang2013second}, Jiang extended the Farkas formulation for the solution of the modified
biharmonic equation, albeit for different boundary conditions. With minor modifications, their formulation
can also be adopted for the solution of the Dirichlet problem for the oscillatory biharmonic equation
\begin{align}
\Delta (\Delta + k^2) w &= 0 \quad \text{in} \quad \Omega \label{eq:helmbi} \\
w &= f \quad \text{on} \quad \Gamma \label{eq:helmbibc1} \\
\dwdn &=g \quad \text{on} \quad \Gamma \label{eq:helmbibc2} \,,
\end{align}
where $f,g$ are prescribed functions on the boundary.

For two points on the plane, $\xx$ and $\yy$, let $\br = \yy - \xx$. Let the kernels $K_{1}^{F}$, and 
$K_{2}^{F}$ be defined by 
\begin{align}
  \Kfark_1(\xx,\yy) &= \partial_{n_{y}n_{y}n_{y}}\Gbh(\xx,\yy)
  + 3 \partial_{n_{y}\tau_{y}\tau_{y}}\Gbh(\xx,\yy) + k^{2} \partial_{n_{y}}\Gbh(\xx,\yy)\; , \label{eq:kf1} \\
  \Kfark_2(\xx,\yy) &= -\partial_{n_{y}n_{y}}\Gbh(\xx,\yy)
  + \partial_{\tau_{y}\tau_{y}}\Gbh(\xx,\yy) \label{eq:kf2} \; .
\end{align}
More explicitly, we have
\begin{align}
  \Kfark_1(\xx,\yy) &=  -\dfrac{\br \cdot \bn(\yy)}{|\br|^2}  \cdot \left( \frac{3 i}{2} H_0^{(1)}(k|\br|) - \frac{5 i}{2 k|\br|} H_1^{(1)}(k|\br|) + \frac{6}{\pi k^2 |\br|^2} + \frac{1}{2\pi}  \right) +     \\
  & \quad \dfrac{\left(\br \cdot \bn(\yy)\right)^3}{|\br|^4} \cdot \left( 2i H_0^{(1)}(k|\br|) - \frac{7 i}{2 k|\br|} H_1^{(1)}(k|\br|) + \frac{8}{\pi k^2 |\br|^2}  \right) \, , \nonumber \\
  \Kfark_2(\xx,\yy) &=  -\left(\frac{1}{2}-\dfrac{\left(\br \cdot \bn(\yy)\right)^2}{|\br|^2} \right)  \cdot \left( \frac{i}{2} H_0^{(1)}(k|\br|) - \frac{i}{k|\br|} H_1^{(1)}(k|\br|) + \frac{2}{\pi k^2 |\br|^2}  \right)\; .
\end{align}

On enforcing the Dirichlet boundary conditions for
$w$, we obtain the integral equation
\begin{equation}
  \begin{split}
    \begin{pmatrix}
      f(\xx) \\
      g(\xx)
    \end{pmatrix} &= 
    \int_\Gamma
    \begin{pmatrix}
      \Kfark_{11}(\xx,\yy) & \Kfark_{12}(\xx,\yy) \\
      \Kfark_{21}(\xx,\yy) & \Kfark_{22}(\xx,\yy)
    \end{pmatrix}
    \begin{pmatrix}
      \sigma_1(\yy)\\
      \sigma_2(\yy)
    \end{pmatrix}
    \, dS(\yy) \\
    &\quad + \begin{pmatrix}
      1/2 & 0 \\
      -\kappa(\xx) & 1/2
    \end{pmatrix}
    \begin{pmatrix}
      \sigma_1(\xx)\\
      \sigma_2(\xx)
    \end{pmatrix} \; , \label{eq:farkIntEq} 
  \end{split}
\end{equation}
where $\kappa$ is the signed curvature at $\xx \in \Gamma$.
The kernels are given
by $\Kfark_{11} = \Kfark_1$, $\Kfark_{12} = \Kfark_2$,
\begin{align}
  \Kfark_{21}(\xx,\yy) &= \left(\Kfark_1(\xx,\yy)\right )_{n_x} \nonumber \\
  &= \bn(\xx) \cdot \nabla_{\xx} K_{21}^{F}(\xx,\yy) \; \label{eq:farkK21},\\
  \Kfark_{22}(\xx,\yy) &= \left(\Kfark_2(\xx,\yy)\right )_{n_x} \nonumber \\
  &= \bn(\xx) \cdot \nabla_{\xx} K_{22}^{F}(\xx,\yy) \label{eq:farkK22} \; .
\end{align}

The following lemma shows that the integral equation~\cref{eq:farkIntEq} is invertible for smooth domains
whenever $k$ is not a buckling eigenfrequency.

\begin{lem}
Suppose that $\Omega \subset \R^2$ is a simply connected region whose boundary $\Gamma$ is $\cC^{2}$. 
Furthermore, suppose that $k$ is not a buckling eigenfrequency of $\Omega$, i.e., the only solution $w$ to~\cref{eq:buck1,eq:buck2,eq:buck3} is  $w \equiv 0$. Suppose further that $f,g \in \cC^{2}(\Gamma)$. 
Then there exists a unique solution $\sigma_{1},\sigma_{2} \in \cC^{2}$ which satisfies~\cref{eq:farkIntEq}. 
Furthermore, $w(\bx)$ defined via~\cref{eq:farkRep} satisfies the oscillatory biharmonic equation~\cref{eq:helmbi}
along with the boundary conditions $w(\bx)=f(\bx)$ and $\dwdn(\bx) = g(\bx)$ for $\bx \in\Gamma$.
\label{lem:farkinv}
\end{lem}

\begin{proof}
  The proof is a minor modification of the one presented in~\ref{eq:shidongmodbi}
  and is contained in~\cref{sec:farkproof}.
\end{proof}

The kernels $\Kfark_1$ and $\Kfark_2$ are
constructed with the goal that the $\Kfark_{ij}$ are
as smooth as possible. Suppose that the boundary
$\Gamma$ is of class $C^k$. Then, the kernels,
$\Kfark_{ij}(\xx,\yy)$, are $C^{k-2}$ functions on the boundary
for each $\yy \in \Gamma$.
Therefore, on a smooth
boundary, these kernels are smooth. However, on a
domain with a corner, we note 
that the kernel $\Kfark_{21}$
has a singularity with strength $r^{-2}$ (see~\cref{fig:ker21plot}). 
This singularity, in addition to the term in \cref{eq:farkIntEq}
which explicitly involves the
curvature, makes the representation \cref{eq:farkRep}
unstable for domains with high curvature (or
corners).

>>>>

Let $\mathcal{L}_x$ denote a linear differential operator. A fundamental
solution $G(\xx,\yy)$ of $\mathcal{L}_x$ satisfies the equation
$\mathcal{L}_x G(\xx,\yy) = \delta_y(\xx)$ in the distributional sense, i.e.
for sufficiently smooth $f$
\begin{equation}
  \mathcal{L}_x \int_{\R^2} G(\xx,\yy) f(\yy) \, dy = f(\xx) \; .
\end{equation}
We consider here
``free-space'' Green's functions, i.e. fundamental solutions which satisfy
some natural growth or decay conditions as $|\xx-\yy| \to \infty$.
The Green's function of the oscillatory biharmonic equation
\cref{eq:buck1} is given by 
\begin{equation}
  \Gbh(\xx,\yy) = \frac{1}{k^2}
  \left (\frac{1}{2\pi} \log |\xx-\yy| +
  \frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \right ) \, , 
\end{equation}
where $k$ is the Helmholtz paramater in the oscillatory biharmonic equation,
and $H_{0}^{1}(r)$ is the Hankel function of the first kind of order zero.
Note that this is a scaled difference of the Green's function for
Laplace, i.e.

@@
